Short-range correlation physics at low RG resolution

### 2020-08-18 Notes/Operator_evolution/referee_response.tex

% Response to referee comments.

\documentclass[preprintnumbers,floatfix,aps,prc,preprint,nofootinbib]{revtex4-1}

\usepackage{physics} % For bra-ket notation

\newcommand{\ataq}{a^{\dagger}_q a_q}
\newcommand{\PP}{\ensuremath{P\mbox{--}P}}
\newcommand{\PQ}{\ensuremath{P\mbox{--}Q}}
\newcommand{\QP}{\ensuremath{Q\mbox{--}P}}
\newcommand{\QQ}{\ensuremath{Q\mbox{--}Q}}
\newcommand{\Klo}{K_{\text{lo}}}
\newcommand{\Khi}{K_{\text{hi}}}
\newcommand{\LambdaBD}{{\Lambda_{\text{BD}}}}

\begin{document}

We thank the referee for her/his careful reading of our paper and for the questions and suggestions. We address each of the points in turn.

\begin{enumerate}

    \item ``The results of section II nicely demonstrate the benefits of the Magnus expansion. I would find it useful if the authors could add some brief discussion on the computational cost of the Magnus expansion (as a function of $k_{\rm max}$) compared to the traditional SRG evolution for a given choice of tolerance for the ODE solver.''

    For the range of applications in the paper, the computational cost is roughly comparable to the conventional free-space SRG.
    While the evaluation of the RHS of the Magnus formulation Eq.~(7) is more complicated with increasing $k_{\rm max}$, it allows the use of a simple ODE solver such as the first-order Euler method, which takes fewer evaluations of the RHS than a typical adaptive high-order solver required to solve Eq.~(2) to comparable precision.
    We've added a sentence to this effect in section II B (on page 5).
    To reiterate what we say in the paper, the computational benefits of the Magnus approach are particularly for the in-medium SRG.

    \item ``At the beginning of section III the authors should add a statement to clarify to what extend the studies in this section are connected to the previous section. First, I assumed that the SRG studies would also be based on the Magnus expansion, this time applied to the high-order NN interaction. However, it seems that these two sections are not immediately related. Is there a reason why the Magnus method was not used for the calculations? Is the agreement between the Magnus method and the traditional evolution expected to be comparable to the results shown in section II?''

    We use the SRG for sections III-IV because the Magnus implementation and typical SRG approach are indistinguishable with soft potentials, unlike the case in section II where we tested high-cutoff potentials, which are intrinsically hard. We added a statement clarifying this at the end of the first paragraph in section III.

    \item ``I find the results shown if Fig.~16 somewhat confusing, especially those in the left three panels. What is the significance of the apparent oscillatory behavior? In the main text these features are attributed to the 'dependence on the mesh'. What does that mean? Is it then even useful to study these matrix elements? I would naively expect that the P-P part in the left panel should be a smooth function without the visible oscillatory artefacts. It would be helpful for me if the authors could comment on that in some more detail.''

    The oscillations seen in Fig.~16 are dependent on the bare $r^2$ operator, which depends on its regularization (which is what we meant by dependent on the mesh) as illustrated in Figs.~14 and 15.
    It is useful to visualize these results, which illustrate factorization and in particular the form of the \QQ\ part.
    We can see the origin of the oscillations by using factorization of the SRG transformation, where $U(k,q) \approx \Klo(k) \Khi(q)$ and from Ref.~[26], $U(k'',k') \approx Z(\LambdaBD) \delta(k''-k')$ where $k,k',k'' < \LambdaBD \ll q$.
    We can take the $\QP$ contribution to $Pr^2P$ as an example. It has the approximate form 
    \begin{align}
       \Klo(k) &\int_{\LambdaBD}^{\infty} dq\, \int_0^{\LambdaBD} dk''\, \Khi(q) \mel{q}{r^2}{k''} Z(\LambdaBD) \delta(k''-k')
       \notag \\
       & \approx 
       \Klo(k) \int_{\LambdaBD}^{\infty} dq\, \Khi(q) \mel{q}{r^2}{k'} Z(\LambdaBD)
    \end{align}
    The $k$-dependence drops out (to leading order $\Klo(k)$ is a constant), and the $k'$-dependence is oscillatory due to the form of the bare $r^2$.
    This explains the vertical stripes in the middle two panels of Fig.~16 being nearly constant within each stripe.
    The oscillatory artefacts are based on setting a sharp cutoff in coordinate space (i.e., an unregulated coordinate-space operator, see Fig.~14 for momentum-space matrix elements).
    Indeed, we can regulate the coordinate-space $r^2$ operator and smooth out the oscillations as in Fig. 15, which keeps a similar induced form in the $\QQ$ panel.
    The oscillations in the \PP\ panel are emphasized because we subtract out the initial $P r^2 P$ operator in Fig.~16;
    we have added a note about this in the caption of Fig.~16.
    At the end of section IV A we added some discussion on the non-$\QQ$ panels, pointing to section IV C for more detail.
    In section IV C, we added further explanation of the terms in Eq.~(15) and Fig.~16 using factorization.

    \item ``It seems that several of the presented results in section IV are directly related to the discussions and results presented in Ref.~[26]. It would be helpful if the authors could add a paragraph in which they summarize the new aspects and insights, as well as differences and similarities of the discussed results compared to Ref.~[26].''

    Our results in Sec.~IV heavily build on the results of Ref.~[26], but extend the conclusions of Ref.~[26] with an emphasis on scheme dependence by contrasting several NN potentials, utilizing band- and block-diagonal SRG decoupling, and varying the bare operators through the implementation of smeared delta functions in $\ataq$ or adding a coordinate-space regulator to $r^2$. We highlight the different forms of scheme dependence in several examples throughout the section. 
    We also apply factorization more widely (as for the $r^2$ analysis) and discuss the implications for different knock-out experiments.
    To summarize these new aspects, we added a paragraph to the end of Sec.~IV on how scheme dependence affects our conclusions on SRG operator evolution.

\end{enumerate}

\end{document}

### 2021-09-10 notes/src_physics/src_low_res.tex
Updating notes/src_physics files after updating arXiv submission
Updated figures and corrected typos after iteration with PRC.

\begin{abstract}
Recent experiments have succeeded in isolating processes for which short-range correlation (SRC) physics is dominant and well accounted for by SRC phenomenology.
But an alternative and compelling picture emerges from renormalization group (RG) evolution to low RG resolution.
At high RG resolution, SRCs are identified as components in the nuclear wave function with relative pair momenta greater than the Fermi momentum.
Scale separation results in wave-function factorization that can be exploited with phenomenologies such as the generalized contact formalism or the low-order correlation operator approximation. 
Evolution to lower resolution shifts SRC physics from nuclear structure to the reaction operators without changing the measured observables.
We show how the features of SRC phenomenology manifested at high RG resolution are cleanly identified 
at low RG resolution
using simple two-body operators and local-density approximations with uncorrelated wave functions,
all of which can be systematically generalized.
We verify that the experimental consequences to date follow directly at low resolution from well-established properties of nucleon-nucleon interactions such as the tensor force.
Thus the RG reconciles the contrasting pictures of the same experiment and shows how to get correct results using wave functions without SRC components.
Our demonstration has implications for the analysis of knock-out reactions for which SRC physics is not cleanly isolated.

\end{abstract}

@@

Short-range correlations (SRCs) in atomic nuclei are usually identified as components of the nuclear wave function with nucleon pair momenta well above the Fermi momentum~\cite{Hen:2016kwk}. 
There has long been an apparent need for such SRCs to account for measured cross sections~\cite{Brueckner:1955zzd}, but direct evidence has been nebulous until recent experiments succeeded in cleanly isolating this physics~\cite{Hen:2016kwk,Korover:2014dma,Hen:2014nza,Duer:2018sby,Duer:2018sxh,Schmookler:2019nvf,Cruz-Torres:2020uke,Schmidt:2020kcl,Korover:2020lqf}. 
SRC phenomenologies have been developed that account for the observations, but they seem to be at odds with successful descriptions of nuclear structure such as the shell model that do not feature explicit short-range structure in the nuclear wave function. 
The application of the renormalization group (RG) can make sense of this conflict. 
RG methods are used to analyze critical phenomena in condensed matter and evolve the strong coupling and parton distributions in high-energy quantum chromodynamics (QCD).
Applied to nuclei, the RG shows how SRC physics is manifested differently at varying resolution scales. 
Here we will illustrate how the RG can bridge low- and high-resolution treatments of the same experiment, shedding light on the implications of SRC physics and on long-standing discrepancies between theory and experiment~\cite{Aumann:2020tcq}.

@@

\begin{figure}[tbh]
    \centering
    (a)\includegraphics[width=0.44\columnwidth]{low_rg_resolution_cartoon_1.pdf}~~%
    (b)\includegraphics[width=0.44\columnwidth]{high_rg_resolution_cartoon_1.pdf}
    \caption{Cartoon snapshots of a nucleus at (a) low-RG  and (b) high-RG resolutions. 
    The back-to-back nucleons at high-RG resolution are an SRC pair with small center-of-mass momentum.
    }
    \label{fig:cartoon_nuclei}
\end{figure}

\begin{figure}[tbh]
    \centering
    \includegraphics[width=0.9\columnwidth]{Cartoon_momentum_distribution.pdf}
    \caption{Cartoon of a nuclear single-particle momentum distribution at high RG resolution.
    The region below $\kF$ is a Fermi sea of nucleons.
    Above $\kF$ the distribution is dominated by SRC pairs, with center-of-mass momenta of order $\kF$.
    At low RG resolution, the upper branch is greatly suppressed.
    }
    \label{fig:cartoon_mom_dist}
\end{figure}

For visualization, cartoon pictures of these resolutions are shown in Figs.~\ref{fig:cartoon_nuclei} and \ref{fig:cartoon_mom_dist}.
Although only qualitative, such pictures give intuition for interpreting experiments.
In Fig.~\ref{fig:cartoon_nuclei} we visualize configurations that might be sampled at different RG resolutions. Only at high-RG resolution do we find a high-momentum pair.
In practice we associate the high-resolution picture with local phenomenological Hamiltonians, in particular the Argonne V18 (AV18) potential~\cite{Wiringa:1994wb} and its associated three-body forces, which have been featured in most of the recent analyses of SRC experiments. 
AV18 describes nucleon-nucleon observables within experimental accuracy up to the inelastic threshold and the elastic part of the cross section reasonably well up to much higher energies.
The AV18 momentum distributions in nuclei extend well above 4\,fm$^{-1}$ or 800\,MeV (in units where $\hbar = c = 1$), as caricatured in Fig.~\ref{fig:cartoon_mom_dist} (see also the figures in Sec.~\ref{sec:low_resolution}).
For concreteness we associate the low-resolution picture with the shell model (configuration interaction), either in its phenomenological form or derived from ab initio methods such as coupled cluster or the in-medium similarity RG.
In this picture, nucleon momentum distributions die off rapidly above $\kF$.

@@

\end{enumerate}
These effects are all consistent with two phenomenological approaches. 
The Generalized Contact Formalism (GCF) model uses a factorization ansatz for the nuclear many-body wavefunction applied to these high-momentum transfer processes~\cite{Weiss:2015mba,Weiss:2016obx,Weiss:2018tbu,Weiss:2021zyb}. 
The low-order correlation operator approximation (LCA)~\cite{Vanhalst:2014cqa,Ryckebusch:2018rct,Ryckebusch:2019oya} is used to compute how SRCs affect nuclear momentum distributions; it does so by shifting the SRC physics from the wave function to a correlation operator, which is in turn parametrized
(the LCA will be discussed further in Sec.~\ref{sec:low_resolution}).

The GCF embodies the pictures in Fig.~\ref{fig:cartoon_nuclei}(b) and Fig.~\ref{fig:cartoon_mom_dist}, with the nuclear wavefunction relevant for SRC physics events represented as a product of a short-distance piece that multiplies a mean-field piece.
When applied to the distribution of pair momenta in coordinate and momentum space (meaning the probability to find two nucleons within a distance $r$ or having a relative momentum $q$), these quantities take the factorized form:
%
\begin{align}
     \rho^A_{{\rm\scriptscriptstyle GCF},\alpha}(r)  &= C_{A}^{NN,\alpha} |\phi^\alpha_{NN}(r)|^2 \;,
     \label{eq:GCF_coordinate}
    \\
    n^A_{{\rm\scriptscriptstyle GCF},\alpha}(q) &= C_{A}^{NN,\alpha} |\widetilde \phi^\alpha_{NN}(q)|^2
    \;. \label{eq:GCF_momentum}
\end{align}
%
Here $A$ is the nucleon number (protons plus neutrons) and $\alpha$ is a spin index.
The common coefficient is called a ``contact'' in analogy to the quantity defined in cold atom physics.
The two-body functions are extracted from the short-distance (high-momentum) dependence of the zero-energy solutions to the Sch\"odinger equation for a given NN interaction.%
\footnote{It would be misleading to call them wave functions, as pointed out originally in Ref.~\cite{Brueckner:1955zzd}.}

@@

\begin{figure}[tbh]
  \centering
  \includegraphics[width=0.9\columnwidth]{AV18_percentages_deuteron_band.png}
  \caption{Percentage contributions from each channel to the matrix element
  of the momentum distribution in the deuteron with $q=2$, $3$, and $4\,\fmi$ for AV18. We compare unevolved and SRG-evolved (both wave function and operator, so the net matrix element is unchanged) results where $\lambda=1.35\,\fmi$.}
  \label{fig:barcharts_AV18}
\end{figure}

\begin{figure}[tbh]
  \centering
  \includegraphics[width=0.9\columnwidth]{Gezerlis_percentages_deuteron_band.png}
  \caption{Same as Fig.~\ref{fig:barcharts_AV18} but for the Gezerlis \NNLO\ potential.}
  \label{fig:barcharts_Gezerlis}
\end{figure}

One might have expected that the dominance of the tensor means that the operator will be sensitive to $^3\mbox{D}_1$ admixtures in the ground state even when evolved.
This is not the case, as illustrated in Figs.~\ref{fig:barcharts_AV18} and \ref{fig:barcharts_Gezerlis}. 
The high-resolution results ($\lambda = \infty$) do come dominantly from the $\tripletD$--$\tripletD$ part for momenta $q$ where the tensor force dominates while it is more evenly split with $\tripletS$--$\tripletS$ (in a scheme/scale dependent way) at higher momenta.
But the low-resolution results ($\lambda = 1.35\,\fmi$ here) are always heavily dominated by the $\tripletS$--$\tripletS$ part.
This implies that simple ground-state wave functions with local density approximations (LDAs) will work quite well at low RG resolution, as illustrated in the next section.

@@

In this section we use second quantization to more precisely characterize the SRC physics and show how simplified calculations (e.g., using LDAs) are meaningful.
The details of how factorization works in second-quantized form have already been worked out in Ref.~\cite{Bogner:2012zm}, which we build on for the current discussion.
The results presented here are only a sampling of what is possible, using the crudest approximation.
Nevertheless, they suffice to show that low-RG resolution calculations can reproduce all the features of SRG phenomenology listed in Sec.~\ref{sec:introduction}.
In future work we will explore more accurate approximations and calibrate the theoretical uncertainties~\cite{Tropiano:2021prep}.

@@

At low resolution there are various options for calculating SRC physics as manifested in the experiments described in Sec.~\ref{sec:introduction}.
By softening the Hamiltonian, the nuclear ground-state wave functions become less correlated, more amenable to many-body perturbation theory, and more universal in nature.
Furthermore, for operators evaluated at the highest momenta, the details of long-range correlations should become less important.
Indeed, the physics is focused on short distances, which suggests that an LDA should work well, particularly as the unitary transformations in the factorization region ($\delUtilde^{(2)}_\lambda(\kvec,\kvec')$ with $|\kvec| < \lambda$ and $|\kvec'|\gg \lambda$) are weakly dependent on the low momentum~\cite{Tropiano:2020zwb}.

@@

We demonstrate the LDA for proton momentum distributions using the AV18 potential in Fig.~\ref{fig:proton_momentum_distributions}, for which we can compare to quantum Monte Carlo calculations~\cite{Wiringa:2013ala,Lonardoni:2017egu,Wiringa:single_distributions}.
We use proton and neutron densities generated from the SLy4 Skyrme interaction using the HFBRAD code~\cite{Bennaceur:2005mx}.
We expect the approximations to be valid at least for momenta above the gray-shaded regions and indeed the agreement is quite reasonable, particularly at the highest momenta.
(We exclude the predictions below $q = 0.6\,\fmi$ because of the poor approximation; better treatments will be presented in Ref.~\cite{Tropiano:2021prep}.)
It is evident that the high momentum tails are very similar across the nuclei; this is the manifestation of the universal behavior of these distributions.
From Table~\ref{tab:partial_wave_contributions} we see evidence that the s-waves dominate the contributions to these momentum distributions.

\begin{table}[htb]
	\caption{Percentage contributions from s-waves and selected p-waves to proton momentum distributions for $q>2\,\fmi$. We show $N_X/N$ with $N \sim \int_2^{\infty} dq\, q^2 n_{\lambda}^A(q)$, where $n_{\lambda}^A(q)$ is the distribution for nucleus $A$ with evolved operator at $\lambda = 1.35\,\fmi$ and $X$ denotes using only one partial wave.}
	\label{tab:partial_wave_contributions}
	\begin{ruledtabular}
		\begin{tabular}{c|ccccc}
		   Nucleus & $^{1}$S$_0$ [\%] & $^{3}$S$_1$ [\%] & $^{3}$P$_0$ [\%] & $^{1}$P$_1$ [\%] & $^{3}$P$_1$ [\%]
		   \mystrut\\
			\colrule
      			$^{12}$C & $15.5$ & $78.9$ & $1.0$ &
      			$0.8$ & $3.8$
                    \mystrut\\
      			$^{16}$O & $15.5$ & $79.1$ & $1.0$ &
      			$0.8$ & $3.6$
                    \mystrut\\
      			$^{40}$Ca & $15.4$ & $78.7$ & $1.1$ &
      			$0.9$ & $4.0$
                    \mystrut\\
      			$^{48}$Ca & $15.4$ & $78.5$ & $1.1$ &
      			$0.9$ & $4.1$
                    \mystrut\\
      			$^{56}$Fe & $15.4$ & $78.4$ & $1.1$ &
      			$0.9$ & $4.2$
                    \mystrut\\
      			$^{208}$Pb & $15.4$ & $78.4$ & $1.1$ &
      			$0.9$ & $4.2$
                    \mystrut
		\end{tabular}
  	\end{ruledtabular}
\end{table}

The high-momentum tail of any momentum distribution will be scale- and scheme-dependent, i.e., dependent on the potential used. 
We can reproduce results at low-RG resolution for any initial potential, but the reaction operators must consistently evolved.
As evident from Fig.~\ref{fig:schematic_SRC_physics}, we can cancel out the scale and scheme dependence in many cases by taking ratios.
We show examples in Figs.~\ref{fig:pp_plus_pn_ratios}, and \ref{fig:pp_to_pn_ratios} for the pair distribution at $Q=0$.
A clarifying feature of the LDA is that the theta functions defining the Fermi sea imply simple consequences for the pair distribution when $Q=0$.
In particular, the proton and neutron Fermi spheres overlap, so when $N \geq Z$ the product of the theta functions  is unity for the proton Fermi momentum below $\kF^p$, independent of $\kF^n$.

@@

In Fig.~\ref{fig:pp_plus_pn_ratios} we show ratios of the pp + pn to nn + np pair momentum distributions at $Q=0$ as a function of $q$ for six nuclei from $^{12}$C to $^{208}$Pb.
We restrict the range of the plot to the region where SRC physics is expected to dominate.
The ratio is equal to one for $N=Z$ nuclei at all $q$, as expected from factorization as illustrated in Fig.~\ref{fig:schematic_SRC_physics} if the nucleus in numerator and denominator are the same.
Even though there are no high-momentum nucleons in the low-resolution wave function, the result is consistent with phenomenology that the proton and neutron high-momentum distributions should be about the same.
In the region of np dominance near $q=2\,\fmi$ (400\,MeV), this ratio should be unity independent of $N/Z$, while there is $N/Z$ dependence away from this region because there are more nn than pp pairs.

>>>>

\begin{abstract}
Recent experiments have succeeded in isolating processes for which short-range correlation (SRC) physics is dominant and well accounted for by SRC phenomenology.
But an alternative and compelling picture emerges from renormalization group (RG) evolution to low RG resolution.
At high RG resolution, SRCs are identified as components in the nuclear wave function with relative pair momenta greater than the Fermi momentum.
Scale separation results in wave-function factorization that can be exploited with phenomenologies such as the generalized contact formalism or the low-order correlation operator approximation. 
Evolution to lower resolution shifts SRC physics from nuclear structure to the reaction operators without changing the measured observables.
We show how the features of SRC phenomenology manifested at high RG resolution are cleanly identified 
at low RG resolution
using simple two-body operators and local-density approximations with uncorrelated wave functions,
all of which can be systematically generalized.
We verify that the experimental consequences to date follow directly at low resolution from well-established properties of nucleon-nucleon interactions such as the tensor force.
Thus the RG reconciles the contrasting pictures of the same experiment and shows how to get correct results using wave functions without SRC components.
Our demonstration has implications for the analysis of knock-out reactions for which SRC physics is not cleanly isolated.
\end{abstract}

@@

Short-range correlations (SRCs) in atomic nuclei are usually identified as components of the nuclear wave function with nucleon pair momenta well above the Fermi momentum~\cite{Hen:2016kwk}. 
There has long been an apparent need for such SRCs to account for measured cross sections~\cite{Brueckner:1955zzd}, but direct evidence has been nebulous until recent experiments succeeded in cleanly isolating this physics~\cite{Hen:2016kwk,Korover:2014dma,Hen:2014nza,Duer:2018sby,Duer:2018sxh,Schmookler:2019nvf,Cruz-Torres:2020uke,Schmidt:2020kcl,CLAS:2020rue}. 
SRC phenomenologies have been developed that account for the observations, but they seem to be at odds with successful descriptions of nuclear structure such as the shell model that do not feature explicit short-range structure in the nuclear wave function. 
The application of the renormalization group (RG) can make sense of this conflict. 
RG methods are used to analyze critical phenomena in condensed matter~\cite{Amit:2005} and evolve the strong coupling and parton distributions in high-energy quantum chromodynamics (QCD)~\cite{RevModPhys.67.157,peskin1995introduction}.
Applied to nuclei, the RG shows how SRC physics is manifested differently at varying resolution scales. 
Here we will illustrate how the RG can bridge low- and high-resolution treatments of the same experiment, shedding light on the implications of SRC physics and on long-standing discrepancies between theory and experiment~\cite{Aumann:2020tcq}.

@@

\begin{figure}[tbh]
    \centering
    (a)\includegraphics[width=0.43\columnwidth]{low_RG_resolution_cartoon_1_new.pdf}~~%
    (b)\includegraphics[width=0.44\columnwidth]{high_RG_resolution_cartoon_1_new2.pdf}
    \caption{Cartoon snapshots of a nucleus at (a) low-RG  and (b) high-RG resolutions. 
    The back-to-back nucleons at high-RG resolution are an SRC pair with small center-of-mass momentum.
    }
    \label{fig:cartoon_nuclei}
\end{figure}

\begin{figure}[tbh]
    \centering
    \includegraphics[width=0.9\columnwidth]{Cartoon_momentum_distribution_v2.pdf}
    \caption{Cartoon of a nuclear single-particle momentum distribution at high RG resolution.
    The region below $\kF$ is a Fermi sea of nucleons.
    Above $\kF$ the distribution is dominated by SRC pairs, with center-of-mass momenta of order $\kF$.
    At low RG resolution, the upper branch is greatly suppressed.
    }
    \label{fig:cartoon_mom_dist}
\end{figure}

For visualization, cartoon pictures of these resolutions are shown in Figs.~\ref{fig:cartoon_nuclei} and \ref{fig:cartoon_mom_dist}.
Although only qualitative, such pictures give intuition for interpreting experiments.
In Fig.~\ref{fig:cartoon_nuclei} we visualize configurations that might be sampled at different RG resolutions. Only at high-RG resolution do we find a high-momentum pair.
In practice we associate the high-resolution picture with local phenomenological Hamiltonians, in particular the Argonne V18 (AV18) potential~\cite{Wiringa:1994wb} and its associated three-body forces~\cite{Carlson:1983kq,Pieper:2001ap}, which have been featured in most of the recent analyses of SRC experiments~\cite{Hen:2016kwk}. 
AV18 describes nucleon-nucleon observables within experimental accuracy up to the inelastic threshold and the elastic part of the cross section reasonably well up to much higher energies.
The AV18 momentum distributions in nuclei extend well above 4\,fm$^{-1}$ or 800\,MeV (in units where $\hbar = c = 1$), as caricatured in Fig.~\ref{fig:cartoon_mom_dist} (see also the figures in Sec.~\ref{sec:low_resolution}).
For concreteness we associate the low-resolution picture with the shell model (configuration interaction), either in its phenomenological form~\cite{Heyde:2004zz,Ring:2005} or derived from ab initio methods such as coupled cluster or the in-medium similarity RG~\cite{Hergert:2020bxy}.
In this picture, nucleon momentum distributions die off rapidly above $\kF$.

@@

\end{enumerate}
These effects are all consistent with two phenomenological approaches. 
The Generalized Contact Formalism (GCF) model uses a factorization ansatz for the nuclear many-body wavefunction applied to these high-momentum transfer processes~\cite{Weiss:2015mba,Alvioli:2016wwp,Weiss:2016obx,Weiss:2018tbu,Weiss:2021zyb}. 
The low-order correlation operator approximation (LCA)~\cite{Vanhalst:2014cqa,Ryckebusch:2018rct,Ryckebusch:2019oya} is used to compute how SRCs affect nuclear momentum distributions; it does so by shifting the SRC physics from the wave function to a correlation operator, which is in turn parametrized
(the LCA will be discussed further in Sec.~\ref{sec:low_resolution}).

The GCF embodies the pictures in Fig.~\ref{fig:cartoon_nuclei}(b) and Fig.~\ref{fig:cartoon_mom_dist}, with the nuclear wavefunction relevant for SRC physics events represented as a product of a short-distance piece that multiplies a mean-field piece.
When applied to the distribution of pair momenta in coordinate and momentum space (meaning the probability to find two nucleons within a distance $r$ or having a relative momentum $q$), these quantities take the factorized form:
%
\begin{align}
     \rho^A_{{\rm\scriptscriptstyle GCF},\alpha}(r)  &= C_{A}^{NN,\alpha} |\phi^\alpha_{NN}(r)|^2 \;,
     \label{eq:GCF_coordinate}
    \\
    n^A_{{\rm\scriptscriptstyle GCF},\alpha}(q) &= C_{A}^{NN,\alpha} |\widetilde \phi^\alpha_{NN}(q)|^2
    \;. \label{eq:GCF_momentum}
\end{align}
%
Here $A$ is the nucleon number (protons plus neutrons) and $\alpha$ is a spin index.
The common coefficient is called a ``contact'' in analogy to the quantity defined in cold atom physics~\cite{Tan2008contact1,Tan2008contact2}.
The two-body functions are extracted from the short-distance (high-momentum) dependence of the zero-energy solutions to the Schr\"odinger equation for a given NN interaction.%
\footnote{It would be misleading to call them wave functions, as pointed out originally in Ref.~\cite{Brueckner:1955zzd}.}

@@

\begin{figure}[tbh]
  \centering
  \includegraphics[width=0.9\columnwidth]{AV18_Gezerlis_percentages_deuteron_band.png}
  \caption{Percentage contributions from each channel to the matrix element
  of the momentum distribution in the deuteron with $q=2$, $3$, and $4\,\fmi$ for AV18 (left bar at each $\lambda$ value) and Gezerlis \NNLO\ (right bar) potentials. We compare unevolved and SRG-evolved (both wave function and operator, so the net matrix element is unchanged) results where $\lambda=1.35\,\fmi$.}
  \label{fig:barchart_deuteron}
\end{figure}

% \begin{figure}[tbh]
%   \centering
%   \includegraphics[width=0.9\columnwidth]{Gezerlis_percentages_deuteron_band.png}
%   \caption{Same as Fig.~\ref{fig:barcharts_AV18} but for the Gezerlis \NNLO\ potential.}
%   \label{fig:barcharts_Gezerlis}
% \end{figure}

One might have expected that the dominance of the tensor means that the operator will be sensitive to $^3\mbox{D}_1$ admixtures in the ground state even when evolved.
This is not the case, as illustrated in Fig.~\ref{fig:barchart_deuteron}. 
The high-resolution results ($\lambda = \infty$) do come dominantly from the $\tripletD$--$\tripletD$ part for momenta $q$ where the tensor force dominates while it is more evenly split with $\tripletS$--$\tripletS$ (in a scheme/scale dependent way) at higher momenta.
But the low-resolution results ($\lambda = 1.35\,\fmi$ here) are always heavily dominated by the $\tripletS$--$\tripletS$ part.
Note that the same pattern holds for both the phenomenological AV18 potential and the Gezerlis \NNLO\ chiral EFT potential.
This S-wave dominance implies that simple ground-state wave functions with local density approximations (LDAs) will work quite well at low RG resolution, as verified in the next section.

@@

In this section we use second quantization to more precisely characterize the SRC physics and show how simplified calculations (e.g., using LDAs) are meaningful.
The details of how factorization works in second-quantized form have already been worked out in Ref.~\cite{Bogner:2012zm}, which we build on for the current discussion.
The results presented here are only a sampling of what is possible, using the crudest approximation.
Nevertheless, they suffice to show that low-RG resolution calculations can reproduce all the features of SRC phenomenology listed in Sec.~\ref{sec:introduction}.
In future work we will explore more accurate approximations and calibrate the theoretical uncertainties~\cite{Tropiano:2021prep}.

@@

At low resolution there are various options for calculating SRC physics as manifested in the experiments described in Sec.~\ref{sec:introduction}.
By softening the Hamiltonian, the nuclear ground-state wave functions become less correlated, more amenable to many-body perturbation theory, and more universal in nature.
Furthermore, for operators evaluated at the highest momenta, the details of long-range correlations should become less important.
Indeed, the physics is focused on short distances, which suggests that an LDA should work well, particularly as the unitary transformations in the factorization region ($\delUtilde^{(2)}_\lambda(\kvec,\kvec')$ with $|\kvec| < \lambda$ and $|\kvec'|\gg \lambda$) are weakly dependent on the low momentum~\cite{Tropiano:2020zwb}.
Figures included as Supplemental Material [URL will be inserted by publisher] illustrate the factorization of the $\delUtilde\delUtilde^{\dagger}$ term in Eq.~\eqref{eq:pair_mom_dist_op}.

@@

We demonstrate the LDA for proton momentum distributions using the AV18 potential in Fig.~\ref{fig:proton_momentum_distributions}, for which we can compare to quantum Monte Carlo calculations~\cite{Wiringa:2013ala,Lonardoni:2017egu,Wiringa:single_distributions}.
We use proton and neutron densities generated from the SLy4 Skyrme functional~\cite{Chabanat:1997un} using the HFBRAD code~\cite{Bennaceur:2005mx}.
We expect the approximations to be valid at least for momenta above the gray-shaded regions and indeed the agreement is quite reasonable, particularly at the highest momenta.
(We exclude the predictions below $q = 0.6\,\fmi$ because of the poor approximation; better treatments will be presented in Ref.~\cite{Tropiano:2021prep}.)
It is evident that the high momentum tails are very similar across the nuclei; this is the manifestation of the universal behavior of these distributions.
There is a clear signature of the sharp cutoff caused by the $\theta$ functions in Eq.~\eqref{eq:snmd_partial_waves} at momenta near the Fermi momentum $\kF$.
We expect a smoother distribution with higher-order contributions to the DME as well as from long-range correlations.
From Table~\ref{tab:partial_wave_contributions} we see evidence that the s-waves dominate the contributions to these momentum distributions.

\begin{table}[htb]
	\caption{Percentage contributions from s-waves and selected p-waves to proton momentum distributions for $q>2\,\fmi$. We show $N_X/N$ with $N \sim \int_2^{\infty} dq\, q^2 n_{\lambda}^A(q)$, where $n_{\lambda}^A(q)$ is the proton distribution for nucleus $A$ with evolved operator at $\lambda = 1.35\,\fmi$ and $X$ denotes using only one partial wave.}
	\label{tab:partial_wave_contributions}
	\begin{ruledtabular}
		\begin{tabular}{c|ccccc}
		   Nucleus & $^{1}$S$_0$ [\%] & $^{3}$S$_1$ [\%] & $^{3}$P$_0$ [\%] & $^{1}$P$_1$ [\%] & $^{3}$P$_1$ [\%]
		   \mystrut\\
			\colrule
      			$^{12}$C & $15.5$ & $78.9$ & $1.0$ &
      			$0.8$ & $3.8$
                    \mystrut\\
      			$^{16}$O & $15.5$ & $79.1$ & $1.0$ &
      			$0.8$ & $3.6$
                    \mystrut\\
      			$^{40}$Ca & $15.4$ & $78.7$ & $1.1$ &
      			$0.9$ & $4.0$
                    \mystrut\\
      			$^{48}$Ca & $15.4$ & $78.5$ & $1.1$ &
      			$0.9$ & $4.1$
                    \mystrut\\
      			$^{56}$Fe & $15.4$ & $78.4$ & $1.1$ &
      			$0.9$ & $4.2$
                    \mystrut\\
      			$^{208}$Pb & $15.4$ & $78.4$ & $1.1$ &
      			$0.9$ & $4.2$
                    \mystrut
		\end{tabular}
  	\end{ruledtabular}
\end{table}

The high-momentum tail of any momentum distribution will be scale- and scheme-dependent, i.e., dependent on the potential used. 
We can reproduce results at low-RG resolution for any initial potential, but the reaction operators must be consistently evolved.
As evident from Fig.~\ref{fig:schematic_SRC_physics}, we can cancel out the scale and scheme dependence in many cases by taking ratios.
We show examples in Figs.~\ref{fig:pp_plus_pn_ratios}, and \ref{fig:pp_to_pn_ratios} for the pair distribution at $Q=0$.
A clarifying feature of the LDA is that the theta functions defining the Fermi sea imply simple consequences for the pair distribution when $Q=0$.
In particular, the proton and neutron Fermi spheres overlap, so when $N \geq Z$ the product of the theta functions  is unity for the proton Fermi momentum below $\kF^p$, independent of $\kF^n$.

@@

In Fig.~\ref{fig:pp_plus_pn_ratios} we show ratios of the pp + pn to nn + np pair momentum distributions at $Q=0$ as a function of $q$ for six nuclei from $^{12}$C to $^{208}$Pb.
We restrict the range of the plot to the region where SRC physics is expected to dominate.
The ratio is equal to one for $N=Z$ nuclei at all $q$, as expected from factorization as illustrated in Fig.~\ref{fig:schematic_SRC_physics} if the nucleus in numerator and denominator are the same.
Even though there are no high-momentum nucleons in the low-resolution wave function, the result is consistent with phenomenology that the proton and neutron high-momentum distributions should be about the same~\cite{Hen:2016kwk}.
In the region of np dominance near $q=2\,\fmi$ (400\,MeV), this ratio should be unity independent of $N/Z$, while there is $N/Z$ dependence away from this region because there are more nn than pp pairs.

In Fig.~\ref{fig:pp_to_pn_ratios} we show ratios of the pp to pn pair distributions as a function of $q$ for the same nuclei as in Fig.~\ref{fig:pp_plus_pn_ratios}.
The ratio dips down to essentially zero just before 2\,\fmi and then rises to about 0.15 at 4\,\fmi.
This trend is also observed in the GCF calculations~\cite{Cruz-Torres:2019fum,CLAS:2020rue} and reflects the transition from the dominant effect of the tensor force toward the scalar limit.
In the low-resolution calculations here, this dependence on $q$ is purely from the high-momentum part of the factorized unitary transformations.
The ratio shows no dependence on $N/Z$ because the pair momentum distribution at high-$q$ and $Q=0$ is restricted entirely by the proton Fermi sphere in the $\delUtilde\delUtilde^{\dagger}$ term of Eq.~\eqref{eq:pmd_partial_waves} for $N \ge Z$ nuclei.
%smaller of the proton or neutron Fermi spheres in the $\delUtilde\delUtilde^{\dagger}$ term of Eq.~\eqref{eq:pmd_partial_waves}.
%For $N \ge Z$ both the numerator and denominator are constrained by the proton Fermi sphere which cancels in the ratio.
